SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-09-07 maintex.tex
various edits in abs and intro

% citations
\defcitealias{Fu18}{Paper I}

@@

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 220 interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25$\pm$0.1 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of about $\sim$0.1$\pm$0.1 dex. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxy is matched to a set of control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. We also find that that the central sSFR is enhanced by smaller mass ratios and closer projected separations; however, the gradients of the profiles within 1.0 $R$/\reff\ are unchanged by the parameters.
\end{abstract}

@@

%%%Galaxy Evolution%%%
In the $\Lambda$CDM model, galaxy evolution is a hierarchical process. In this model, massive galaxies are the product of several past merger events of smaller galaxies. In fact, cosmological hydrodynamical simulations have shown that repeated merger events may be responsible for as much as $\sim$60\% of stellar mass in massive galaxies like M87 (e.g. \citet{Rodriguez-Gomez:2016}; \citet{Pillepich:2018}). As the galaxies undergo the merging process, the gas within the galaxies are subject to gravitational torques which perturb the morphology of the galaxies.

@@

%%%Observations%%%
Interaction induced star formation was first seen observationally in the bluer colors of peculiar galaxies in \citet{Larson:1978}. This observation has also been shown in more recent works using the single-fiber spectroscopic survey, SDSS (Sloan Digital Sky Survey) \citet{Ellison:2008, Li:2008, Scudder:2012, Patton:2013, Bustamante:2020}. From these previous works it has been shown that the strength of the star formation enhancement in the centers of paired galaxies is dependent on the stellar mass of the pairs \citet{Li:2008}, the projected separation between the pairs \citet{Ellison:2008, Li:2008, Scudder:2012}, and the mass ratio between the pairs \citet{Ellison:2008}.

%%%Integral Field Spectroscopy%%%
The previous mentioned works using SDSS were restricted to studying the centers of the paired galaxies through a 2\arcsec\ diameter optical fiber. With the recent large integral field spectroscopic (IFS) surveys, interacting galaxies can now be studied with unprecedented spatial detail. These surveys allow astronomers to study the centers of merging galaxies more rigorously since apertures can be set to the physical scale of the galaxies instead of being bound by a fixed sky aperture. These IFS surveys will also allow us to see the extent of the centrally induced star formation and to see how the star formation in the disks of the galaxies are affected.

Astronomers have already begun to take advantage of IFS surveys in galaxy pair studies. \citet{Barrera-Ballesteros:2015} used the CALIFA (Calar Alto Legacy Integral Field Area) survey to study a sample of 103 paired galaxies by varying the size of the aperture through which the \ewha\ is extracted from. In the study, \citet{Barrera-Ballesteros:2015} found a moderate enhancement to the sSFR in the centers of paired galaxies and a moderate suppression to the sSFR in outskirts of the paired galaxies.

\citet{Pan:2019} used the MaNGA survey to study radial profiles of a sample of 205 paired galaxies. The enhancement to the sSFR was shown to be the strongest in the centers of the paired galaxies. This central enhancement linearly fell with increasing galactocentric radii; however, a moderate enhancement to the sSFR remain in the outskirts of the galaxies. \citet{Pan:2019} further studied the paired galaxies as a function of merger stage; from well separated pairs to post-merger galaxies. Across the different merger stages, the sSFR enhancement was greatest in close pairs with tidal features and in post-merger galaxies. This was in agreement with previous hydrodynamical simulations which showed that a burst of star formation is triggered after the first pericenter and as the two galaxies begin to coalesce \citep{Scudder:2012}.

The radial profile of sSFR in post-merger galaxies has also been studied with the MaNGA survey by \citet{Thorp:2019}. The post-merger galaxies were shown to have a strong enhancement to the sSFR in their centers as well as a moderate enhancement in their outskirts.

@@

In our previous work, \citet{Fu:2018}, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit (IFU). We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) \citep{Ahumada:2020} and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in Section \ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in Section \ref{sec:results} we study the radial profiles as a function of stellar mass, projected separation, and the mass ratio, in Section \ref{sec:disc} we compare our work against previous works, and in Section \ref{sec:sum} we summarize the findings of the work.

>>>>

% citations
\defcitealias{Fu:2018}{Paper I}

@@

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 220 star-forming galaxies in close pairs with those of mass-matched isolated galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.3 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy.
\end{abstract}

@@

%%%Galaxy Evolution%%%
In the $\Lambda$CDM model, galaxy evolution is a hierarchical process. In this model, massive galaxies are the product of several past merger events of smaller galaxies. In fact, cosmological hydrodynamical simulations have shown that repeated merger events may be responsible for as much as $\sim$60\% of stellar mass in massive galaxies like M87 \citep[e.g.,][]{Rodriguez-Gomez:2016,Pillepich:2018}. As the galaxies undergo the merging process, the gas within the galaxies are subject to gravitational torques which perturb the morphology of the galaxies.

@@

%%%Observations%%%
Interaction induced star formation was first seen observationally in the bluer colors of peculiar galaxies in \citet{Larson:1978}. This observation has also been shown in more recent works using the single-fiber spectroscopic survey, SDSS (Sloan Digital Sky Survey) \citep{Ellison:2008, Li:2008, Scudder:2012, Patton:2013, Bustamante:2020}. From these previous works it has been shown that the strength of the star formation enhancement in the centers of paired galaxies is dependent on the stellar mass of the pairs \citep{Li:2008}, the projected separation between the pairs \citep{Ellison:2008, Li:2008, Scudder:2012}, and the mass ratio between the pairs \citep{Ellison:2008}.

%%%Integral Field Spectroscopy%%%
The previous mentioned works using SDSS were restricted to studying the centers of the paired galaxies through 2\arcsec\ diameter optical fibers. With the recent large integral field spectroscopic (IFS) surveys, interacting galaxies can now be studied with unprecedented spatial detail. These surveys allow us to study the centers of merging galaxies more rigorously since apertures can be set to the physical scale of the galaxies instead of being bound by a fixed sky aperture. These IFS surveys will also allow us to see the extent of the centrally induced star formation and to see how the star formation in the disks of the galaxies are affected.

Indeed, \citet{Barrera-Ballesteros:2015} used the CALIFA (Calar Alto Legacy Integral Field Area) survey to study a sample of 103 paired galaxies by varying the size of the aperture through which the \ewha\ is extracted from. In the study, \citet{Barrera-Ballesteros:2015} found a moderate enhancement to the sSFR in the centers of paired galaxies and a moderate suppression to the sSFR in outskirts of the paired galaxies. 
%
\citet{Pan:2019} used the MaNGA survey to study radial profiles of a sample of 205 paired galaxies. The enhancement to the sSFR was shown to be the strongest in the centers of the paired galaxies. This central enhancement linearly fell with increasing galactocentric radii; however, a moderate enhancement to the sSFR remain in the outskirts of the galaxies. \citet{Pan:2019} further studied the paired galaxies as a function of merger stage; from well separated pairs to post-merger galaxies. Across the different merger stages, the sSFR enhancement was greatest in close pairs with tidal features and in post-merger galaxies. This was in agreement with previous hydrodynamical simulations which showed that a burst of star formation is triggered after the first pericenter and as the two galaxies begin to coalesce \citep{Scudder:2012}. 
%
The radial profile of sSFR in post-merger galaxies has also been studied with the MaNGA survey by \citet{Thorp:2019}. The post-merger galaxies were shown to have a strong enhancement to the sSFR in their centers as well as a moderate enhancement in their outskirts.

@@

In our previous work, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit \citep[][hereafter \citetalias{Fu:2018}]{Fu:2018}. We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) \citep{Ahumada:2020} and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in Section \ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in Section \ref{sec:results} we study the radial profiles as a function of stellar mass, projected separation, and the mass ratio, in Section \ref{sec:disc} we compare our work against previous works, and in Section \ref{sec:sum} we summarize the findings of the work. 
% cosmology
Throughout we adopt the $\Lambda$CDM cosmology with $\Omega_{\rm m}=0.3$, $\Omega_\Lambda=0.7$, and $h=0.7$.
